---
title: "Demonstrating `clusterProfiler`"
author: "Darach"
date: "`r Sys.Date()`"
---

```{r,libs,cache=F}
library(tidyverse)
library(clusterProfiler)
library(org.Sc.sgd.db)
```

```{r,getGOtermz,cache=T}

# GOSlim annotation table
read_tsv("data/sgd_go_slim_171013.txt",col_names=F,
  comment="!") -> SGDGOSlim

# Full GO annotation table
read_tsv("data/sgd_go_full_171013.txt",col_names=F,
  comment="!") -> SGDGOFull

# Go terms look up table
read_tsv("data/sgd_go_terms_171013.txt",col_names=F,
  comment="!") %>%
  mutate(GOID=str_c("GO:"
    ,str_pad(string=X1,width=7,side="left",pad="0"))
    ) -> SGDGOTerms

# This is a handy table for mapping systematic names to SGD ID
SystematicToSGDIDmapper <- SGDGOSlim%>%
  dplyr::select(X1,X3)%>%
  dplyr::rename(Systematic=X1,SGDID=X3)%>%
  unique()
```


Here's the data. It's from the Barseq after FACS after FISH, aka
the Great Red Herring Fishing Expedition of 2015-2017. 

```{r,dataToAnalyze,cache=T}
datar <- read_csv("data/Figure4_Table_BFFfilteredPooledModels.csv",
  comment="#",
  col_names=c("Systematic","Common","PreShift_Mean",
    "PostShift_Mean","Change_Mean","ResidualFromExpected_Mean"))

ggplot(datar)+aes(x=PostShift_Mean)+geom_histogram(bins=50)
```

We could do a categorical test. That's fast. For this, I'm going
to use the `bitr` function from `clusterProfiler`.

First, we make two lists of IDs with this. So that's just separating
on the mean being more than 7.5, and making a query and background
set.

```{r,clusterProfilerSetupCateogries,cache=T}
library(clusterProfiler)

TopPostShift_Entrez <- datar%>% filter(PostShift_Mean>7.5)%>%
  left_join(SystematicToSGDIDmapper,by="Systematic")%>%
  {bitr(.$SGDID,fromType="SGD",toType="ENTREZID"
    ,OrgDb=org.Sc.sgd.db,drop=T)} %>%
  pull(ENTREZID)

Background_Entrez <- datar%>% 
  left_join(SystematicToSGDIDmapper,by="Systematic")%>%
  {bitr(.$SGDID,fromType="SGD",toType="ENTREZID"
    ,OrgDb=org.Sc.sgd.db,drop=T)} %>%
  pull(ENTREZID)

```

Then testing these for each category of GO terms.

```{r,categoricalGO,cache=T}

CategoricalGO <- list()

for (i in c("MF","BP","CC")) {
  CategoricalGO[[i]] <- enrichGO(
      gene=TopPostShift_Entrez,
      universe=Background_Entrez,
      OrgDb=org.Sc.sgd.db,
      ont=i,
      pAdjustMethod="fdr",
      qvalueCutoff=0.20,
      minGSSize=2,maxGSSize=200)@result%>%
    as_tibble()%>%mutate(Type=i)
}

categoricalGOtable <- bind_rows(CategoricalGO) %>% as_tibble %>% 
    # First we bind rows and make it a tibble
  dplyr::select(Type,ID,Description,qvalue,
    Count,geneID) %>%
    # Then we pick the variables we want
  arrange(Type,-qvalue)


write_lines("#
# Supplementary Table
#
# Results of that categorical GO analysis. Put a description
# here please.
#
# Key:
# Type of enrichment , GO term ID , GO term , q-value , 
#   size of GO term count , geneIDs in this",
  path="output_table_a.csv")

categoricalGOtable %>% 
  write_csv(path="output_table_a.csv",append=T)

```

Okay, but what about GSEA?

So, let's sort the names by that variables, and then see who's 
enriched. First, I do one variable, then I do it in a loop.

I set up a parameter of number of permutations. 
`1e6` takes *forever* to run. So test it out with high p.adjust and
low permutes, then move it up to ~`1e6` or so. You need high
permutations to make claims about stuff rare enough to survive
p-value adjustment as significant, so test on low significance and
then let it run in looking for real significance.

```{r,clusterProfilerSetup,cache=T}
nPermutations <- 1e6

# A list for results holding
GSEAResults <- list()
```

```{r,GSEAfromSort,dependson="clusterProfilerSetup",cache=T}

# Doing one

# First, we make a named list of values, and sort by it
this_gene_list <- 
  datar[order(datar[["PostShift_Mean"]],decreasing=T),] %>%
    # First, it is sorted by a variable. 
    # This non-tidy evaluation is used 
    # because I could not find an easy way to evaluate 
    # it programmatically within tidverse's NSE choice.
  left_join(SystematicToSGDIDmapper,by="Systematic")%>%
    # Then it is joined to a mapper table defined above, to map
    # to the SGD ID
  {setNames(.[["PostShift_Mean"]],nm=.[["SGDID"]])}%>%na.omit()%>%c()
    # Then this is used to name the vector of values of 
    # PostShift_Mean as the SGD ID's. This is na.omited, then c'd
    # into a plain named vector.

GSEAResults[["PostShift_Mean"]] <- GSEA(
  geneList=this_gene_list,
    # This is then piped into GSEA function as the gene list
  TERM2GENE=SGDGOSlim[,c("X6","X3")],
    # This provides a mapping of term and gene ID (SGDID)
  TERM2NAME=SGDGOTerms[,c("GOID","X2")],
    # This provides a mapping of term ID to term description
  minGSSize=3,maxGSSize=500,
    # This controls the tested gene set sizes
  pAdjustMethod="fdr",
    # This sets the method of p-value adjusting
  pvalueCutoff=0.1,
    # This sets the p-value cut off
  nPerm=nPermutations,seed="seedy"
    # This sets the permutations for the boostrapped p-value, and
    # the random seed
  )@result %>% as_tibble() 
    # Then it is pulled out and set as a tibble

# Doing multiple variables? One could imagine sorting each way:

for (varz in c("PostShift_Mean","PreShift_Mean","Change_Mean")) {
  message(varz)
  message(" took ... ")
  system.time(
    {
    this_gene_list <- datar[order(datar[[varz]],decreasing=T),] %>%
      left_join(SystematicToSGDIDmapper,by="Systematic")%>%
      {setNames(.[[varz]],nm=.[["SGDID"]])}%>%na.omit()%>%c()
    GSEAResults[[varz]] <- GSEA(
      geneList=this_gene_list,
      TERM2GENE=SGDGOSlim[,c("X6","X3")],
      TERM2NAME=SGDGOTerms[,c("GOID","X2")],
      minGSSize=3,maxGSSize=500,
      pAdjustMethod="fdr",
      pvalueCutoff=0.1,seed="seedy",
      nPerm=nPermutations
      )@result %>% as_tibble() %>% mutate(Variable=varz)
    }
  )
}

```

```{r,printOut,cache=T}

# Here we make it presentable

gseGOtable <- bind_rows(GSEAResults) %>% as_tibble %>% 
    # First we bind rows and make it a tibble
  dplyr::select(Variable,ID,Description,enrichmentScore,p.adjust
    ,setSize,rank) %>%
    # Then we pick the variables we want
  arrange(Variable,-enrichmentScore)

write_lines("#
# Supplementary Table
#
# Results of that gene set enrichment analysis. Put a description
# here please.
#
# Key:
# Variable sorted by , GO term ID , GO term , enrichment score , adjusted p-value , size of GO term set , rank of this set in gene list"
  ,path="output_table.csv")

gseGOtable %>% 
  write_csv(path="output_table.csv",append=T)

```

You can also map this to KEGG terms as well.


```{r}
sessionInfo()
```




